---
title: "Data Import"
output: github_document
---
We are loading the "setup" code chunk. "readxl" is for reading excel files in R. "haven" can be used for reading SAS files in R.
```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters datasets and show the full column specification for this data.

##Quick comment key opt+commend+i##
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
spec(litters_df)

## use janitor::clean_names() to clean up variable names
litter_df = janitor::clean_names(litters_df)
```

## take a look at the data

Printing in the concole.
```{r}
litters_df
```

Check the tail of the dataset. 
```{r}
tail(litter_df)
```

```{r}
skimr::skim(litter_df)
```

## Options to read_csv

Check out `?read_csv()` for more information.

#### Example

litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE, na = c("","NA",".",999), col_types)

- Skip the first 10 rows.
- Show column names just as variables X1, X2, X3...
- Tell R to see "", "NA", ".", and 999 as missing values.



## Other file formats

#### Read in an excel file.
Check out `?read_excel()` for more information.
- range = " " let you only read certain range of data

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```


#### Read in a SAS file.
```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```




