---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

## `select`

#### Keep or delete some columns.

Keep certain or range of columns.

select(dataset_name, varibles you want to keep seperate by comma)
```{r}
select(litters_df, group, litter_number)
```

select(dataset_name, a range of variables)
```{r}
select(litters_df, group, gd0_weight:gd_of_birth )
```

We can specify what we don't want.

select(dataset_name, -varibles you don't want)
```{r}
select(litters_df, -litter_number)
```


#### Renaming solumns

select(dataset_name, new_variable_name = original_name, new_name = old_name)
```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Or, you can use "rename()"
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

#### Select helpers
Check on `?select_helpers`

Select every variables that start with "gd."
```{r}
select(litters_df, starts_with("gd"))
```

We want to see litter_number, but still keep everthing else. litter_number becomes the first variable in the table.
```{r}
select(litters_df, litter_number, everything())
```

Or you can use relocate()
```{r}
relocate(litters_df, litter_number)
```



