---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

# `select`

### Keep or delete some columns.

Keep certain or range of columns.

select(dataset_name, varibles you want to keep seperate by comma)
```{r}
select(litters_df, group, litter_number)
```

select(dataset_name, a range of variables)
```{r}
select(litters_df, group, gd0_weight:gd_of_birth )
```

We can specify what we don't want.

select(dataset_name, -varibles you don't want)
```{r}
select(litters_df, -litter_number)
```


### Renaming solumns

select(dataset_name, new_variable_name = original_name, new_name = old_name)
```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Or, you can use "rename()"
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

### Select helpers
Check on `?select_helpers`

Select every variables that start with "gd."
```{r}
select(litters_df, starts_with("gd"))
```

We want to see litter_number, but still keep everthing else. litter_number becomes the first variable in the table.
```{r}
select(litters_df, litter_number, everything())
```

Or you can use relocate()
```{r}
relocate(litters_df, litter_number)
```



# `filter`
We need to give conditions.
```{r}
filter(litters_df, gd0_weight < 22)
filter(litters_df, gd0_weight >= 22)
```

two "=" sign, "==", means testing the equality
For gf_of_birth that are equal to 20.
```{r}
filter(litters_df, gd_of_birth == 20)
```

For gf_of_birth that are NOT equal to 20, we can use "!="
```{r}
filter(litters_df, gd_of_birth != 20)
```

Or we use a more general way to get gf_of_birth that are NOT equal to 20.
Asking for the samples that not fill in given condition. "!()"
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

Give two conditions. Both of them need to be filled.
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

Select the Con7 group.
```{r}
filter(litters_df, group == "Con7")
```

For groups that are either in Con7 or Mod8. 
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```



# `mutate`
 Example
 - create a new variable wt_gain
 - modify the exsiting variable, change all "group" data into lower case and rewrite them under the "group" variable
```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
```

