---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```


## `pivot_longer`
Load the PULSE data, a SAS file.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format... use `pivot_longer`.

Example

- select the range of variable (in this case, they are multiple bdi measurements)
- put variables names to a new variable "visit"
- we don't want the prefix "bdi_score_"
- put their values to a new variable "bdi"

```{r}
pulse_data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

Rewrite, combine, and extend (to add a mutate)

- We put "id" and "visit" in the front as the 1st and 2nd variables through `relocate()`.
- We also re-code part of the "visit" variable, where "bl" as "00m", through `mutate(visit = recode(visit, "bl" = "00m"))`.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```



## `pivot_wider`

Make up some data.
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean" %>% 
      view()
  )
```



## Binding Rows using `bind_row`

Using the LotR data.

First Step: importing each table


```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all the rows together. Stacking them and cleaning the data.
```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie) %>% 
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  )
```



## Joining Datasets

Import the FAS datasets.

```{r}
pups_df = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  relocate(litter_number) %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)
```

`seperate` is used to separate the "group" column into "dose" and "day_of_tx" colums. The separation took place after the 3rd character. 


Next step, joining them. Merging the litters data into the pups data through `left_joint(into_dataset, other_dataset`.

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
```

Data cleaning

- arrange the order by litter_number
- set litter_number, dose, day_of_tx as the first 3 variables, via `relocate()`.

