---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```


## `pivot_longer`
Load the PULSE data, a SAS file.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format... use `pivot_longer`.

Example

- select the range of variable (in this case, they are multiple bdi measurements)
- put variables names to a new variable "visit"
- we don't want the prefix "bdi_score_"
- put their values to a new variable "bdi"
```{r}
pulse_data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

Rewrite, combine, and extend (to add a mutate)

- We put "id" and "visit" in the front as the 1st and 2nd variables through `relocate()`.
- We also re-code part of the "visit" variable, where "bl" as "00m", through `mutate(visit = recode(visit, "bl" = "00m"))`.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```



